The Role of Variability in Learning Generalization: A Computational Modeling Approach

### 2024-05-27 Presentation/_fig_creation.qmd
slide updates

@@

>>>>

# HTW Results

pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,
  brms,tidybayes, rstanarm,emmeans,broom,bayestestR,
  data.table, stringr, here,conflicted, gt, ggh4x, patchwork, knitr)
#options(brms.backend="cmdstanr",mc.cores=4)
walk(c("brms","dplyr","bayestestR","here"), conflict_prefer_all, quiet = TRUE)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)

walk(c("brms","dplyr","bayestestR","here"), conflict_prefer_all, quiet = TRUE)
walk(c("Display_Functions","deLosh_data","fun_alm","fun_indv_fit","fun_model", "prep_model_data","org_functions"), ~source(here::here(paste0("Functions/", .x, ".R"))))

e1 <- readRDS(here("data/e1_08-21-23.rds")) 
e1Sbjs <- e1 |> group_by(id,condit) |> summarise(n=n())
testE1 <- e1 |> filter(expMode2 == "Test")
nbins=5
trainE1 <-  e1 |> filter(expMode2=="Train") |> group_by(id,condit, vb) |> 
    mutate(Trial_Bin = cut( gt.train, breaks = seq(1, max(gt.train),length.out=nbins+1),include.lowest = TRUE, labels=FALSE)) 
trainE1_max <- trainE1 |> filter(Trial_Bin == nbins, bandInt==800)
trainE1_avg <- trainE1_max |> group_by(id,condit) |> summarise(avg = mean(dist))

condEffects <- function(m,xvar){
  m |> ggplot(aes(x = {{xvar}}, y = .value, color = condit, fill = condit)) + 
  stat_dist_pointinterval() + 
  stat_halfeye(alpha=.1, height=.5) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

}
```

## HTW E1

```{r}
#| label: fig-e1-test-dev
#| fig-cap: Experiment 1 Testing Accuracy. A) Empricial Deviations from target band during testing without feedback stage. B) Conditional effect of condition (Constant vs. Varied) and testing band type (trained bands vs. novel extrapolation bands) on testing accuracy. Error bars represent 95% credible intervals.
#| fig-width: 4
#| fig-height: 4

# pe1td <- testE1 |>  
#   ggplot(aes(x = vb, y = dist,fill=condit)) +
#     stat_summary(geom = "bar", position=position_dodge(), fun = mean) +
#     stat_summary(geom = "errorbar", position=position_dodge(.9), fun.data = mean_se, width = .4, alpha = .7) + 
#   theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
#   labs(x="Band", y="Deviation From Target Band")

pe1td <- testE1 |> ggplot(aes(x = vb, y = dist,fill=condit,
                               col=ifelse(bandType=="Trained","black","lightgray"),
                               size=ifelse(bandType=="Trained",1.0,.5))) +
  stat_bar + 
  theme(legend.title.position="top",
        legend.title = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  labs(x="Band", y="Deviation", fill="Condition") +
     # scale_color_manual(values = c("black" = "black"), guide = "legend") +
  scale_color_manual(values=c("black","lightgray"), 
                     name= "Band Type",
                     labels = c("Trained","Extrapolation"),
                     guide = guide_legend(override.aes = list(fill = NA))
                     ) +
  scale_size_continuous(range = c(0.8, 1.0), guide="none")

bmtd <- readRDS(paste0(here::here("data/model_cache/"), "e1_dist_Cond_Type_RF_2.rds"))
pe1ce <- bmtd |> emmeans( ~condit + bandType) |>
  gather_emmeans_draws() |>
 condEffects(bandType) + labs(y="Deviation", x="Band Type")

#p2 <- (pe1td + pe1ce) + plot_annotation(tag_levels= 'A')
#ggsave(here::here("Assets/figs", "e1_test-dev.png"), p2, width=8, height=4, bg="white")
#p2

e1_vxBMM <- readRDS(paste0(here::here("data/model_cache", "e1_testVxBand_RF_5k.rds")))
pe1vce <- e1_vxBMM |> emmeans( ~condit + bandInt,re_formula=NA, 
                     at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
gather_emmeans_draws() |> 
condEffects(bandInt) +
stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
  geom_abline(slope = 1, intercept = 200, linetype = "dashed", color = "gray", size = .5,alpha=.5) +
scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                   labels = levels(testE1$vb), 
                   limits = c(100, 1200)) + 
scale_y_continuous(expand=expansion(add=100),breaks=round(seq(400,1400,by=200),2)) +
theme(legend.position="none") + 
  #theme(legend.title=element_blank(),legend.position="right") + 
labs(y="Velcoity", x="Band")

fe <- fixef(e1_vxBMM)[,1]
fixed_effect_bandInt <- fixef(e1_vxBMM)[,1]["bandInt"]
fixed_effect_interaction <- fixef(e1_vxBMM)[,1]["conditVaried:bandInt"]

re <- data.frame(ranef(e1_vxBMM, pars = "bandInt")$id[, ,'bandInt']) |> 
  rownames_to_column("id") |> 
  left_join(e1Sbjs,by="id") |>
  mutate(adjust= fixed_effect_bandInt + fixed_effect_interaction*(condit=="Varied"),slope = Estimate + adjust )

pid_den1 <- ggplot(re, aes(x = slope, fill = condit)) + 
  geom_density(alpha=.5) + 
  geom_vline(xintercept = 1, linetype="dashed",alpha=.5) +
  xlim(c(min(re$slope)-.3, max(re$slope)+.3))+
   theme(legend.title=element_blank()) + 
  labs(x="Slope Coefficient",y="Density")

pid_slopes1 <- re |>  mutate(id=reorder(id,slope)) |>
  ggplot(aes(y=id, x=slope,fill=condit,color=condit)) + 
    geom_pointrange(aes(xmin=Q2.5+adjust, xmax=Q97.5+adjust)) + 
  geom_vline(xintercept = 1, linetype="dashed",alpha=.5) +
     theme(legend.title=element_blank(), 
           axis.text.y = element_text(size=6) ) + 
    labs(x="Estimated Slope", y="Participant")  + 
    ggh4x::facet_wrap2(~condit,axes="all",scales="free_y")

pe1td
pe1ce

pe1vce
pid_den1
pid_slopes1

ggsave(here::here("Presentation/assets/htw_e1_test1.png"), pe1td, width = 4.0, height = 3, units = "in", dpi = 300)
ggsave(here::here("Presentation/assets/htw_e1_test2.png"), pe1ce, width = 4.0, height = 3, units = "in", dpi = 300)
ggsave(here::here("Presentation/assets/htw_e1_test3.png"), pe1vce, width = 4.0, height = 3, units = "in", dpi = 300)
ggsave(here::here("Presentation/assets/htw_e1_test4.png"), pid_den1, width = 4.0, height = 3, units = "in", dpi = 300)
ggsave(here::here("Presentation/assets/htw_e1_test5.png"), pid_slopes1, width = 4.0, height = 3, units = "in", dpi = 300)
```

## HTW Experiment 2

```{r}
#| label: fig-e2-test-dev
#| fig-cap: Experiment 2 Testing Accuracy. A) Empricial Deviations from target band during testing without feedback stage. B) Conditional effect of condition (Constant vs. Varied) and testing band type (trained bands vs. novel extrapolation bands) on testing accuracy. Error bars represent 95% credible intervals.
#| fig-width: 5
#| fig-height: 5
e2 <- readRDS(here("data/e2_08-04-23.rds")) 
e2Sbjs <- e2 |> group_by(id,condit) |> summarise(n=n())
testE2 <- e2 |> filter(expMode2 == "Test")
nbins=5
trainE2 <-  e2 |> filter(expMode2=="Train") |> group_by(id,condit, vb) |> 
    mutate(Trial_Bin = cut( gt.train, breaks = seq(1, max(gt.train),length.out=nbins+1),include.lowest = TRUE, labels=FALSE)) 
trainE2_max <- trainE2 |> filter(Trial_Bin == nbins, bandInt==600)

bmtd2 <- readRDS(paste0(here::here("data/model_cache/"), "e2_dist_Cond_Type_RF_2.rds"))

condEffects <- function(m,xvar){
  m |> ggplot(aes(x = {{xvar}}, y = .value, color = condit, fill = condit)) + 
  stat_dist_pointinterval() + 
  stat_halfeye(alpha=.1, height=.5) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))

}

pe2td <- testE2 |> ggplot(aes(x = vb, y = dist,fill=condit,
                               col=ifelse(bandType=="Trained","black","lightgray"),
                               size=ifelse(bandType=="Trained",1.0,.5))) +
  stat_bar + 
  theme(legend.title.position="top",
        legend.title = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  labs(x="Band", y="Deviation", fill="Condition") +
  scale_color_manual(values=c("black","lightgray"), 
                     name= "Band Type",
                     labels = c("Trained","Extrapolation"),
                     guide = guide_legend(override.aes = list(fill = NA))
                     ) +
  scale_size_continuous(range = c(0.8, 1.0), guide="none")

pe2ce <- bmtd2 |> emmeans( ~condit + bandType) |>
  gather_emmeans_draws() |>
 condEffects(bandType) + labs(y="Absolute Deviation From Band", x="Band Type")

e2_vxBMM <- readRDS(paste0(here::here("data/model_cache", "e2_testVxBand_RF_5k.rds")))

pe2vce <- e2_vxBMM |> emmeans( ~condit + bandInt,re_formula=NA, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |> 
  condEffects(bandInt) +
  stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
  geom_abline(slope = 1, intercept = 200, linetype = "dashed", color = "gray", size = .5,alpha=.5) +
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE2$vb), 
                     limits = c(100, 1200)) + 
scale_y_continuous(expand=expansion(add=100),breaks=round(seq(200,1400,by=200),2)) +
  theme(legend.title=element_blank()) + 
  labs(y="Velcoity", x="Band")

fe <- fixef(e2_vxBMM)[,1]
fixed_effect_bandInt <- fixef(e2_vxBMM)[,1]["bandInt"]
fixed_effect_interaction <- fixef(e2_vxBMM)[,1]["conditVaried:bandInt"]

re <- data.frame(ranef(e2_vxBMM, pars = "bandInt")$id[, ,'bandInt']) |> 
  rownames_to_column("id") |> 
  left_join(e2Sbjs,by="id") |>
  mutate(adjust= fixed_effect_bandInt + fixed_effect_interaction*(condit=="Varied"),slope = Estimate + adjust )

pid_den2 <- ggplot(re, aes(x = slope, fill = condit)) + 
  geom_density(alpha=.5) + 
  geom_vline(xintercept = 1, linetype="dashed",alpha=.5) +
  xlim(c(min(re$slope)-.3, max(re$slope)+.3))+
   theme(legend.title=element_blank()) + 
  labs(x="Slope Coefficient",y="Density")

pid_slopes2 <- re |>  mutate(id=reorder(id,slope)) |>
  ggplot(aes(y=id, x=slope,fill=condit,color=condit)) + 
    geom_pointrange(aes(xmin=Q2.5+adjust, xmax=Q97.5+adjust)) + 
  geom_vline(xintercept = 1, linetype="dashed",alpha=.5) +
      theme(legend.title=element_blank(), 
        axis.text.y = element_text(size=6) ) + 
    labs(x="Estimated Slope", y="Participant")  + 
    ggh4x::facet_wrap2(~condit,axes="all",scales="free_y")

pe2td
pe2ce
pe2vce
pid_den2
pid_slopes2

ggsave(here::here("Presentation/assets/htw_e2_test1.png"), pe2td, width = 4.0, height = 3, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e2_test2.png"), pe2ce, width = 4.0, height = 3, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e2_test3.png"), pe2vce, width = 4.0, height = 3, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e2_test4.png"), pid_den2, width = 4.0, height = 3, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e2_test5.png"), pid_slopes2, width = 4.0, height = 3, units = "in", dpi = 600)

```

## HTW Exp 3

```{r}
#| fig-width: 4
#| fig-height: 6
e3 <- readRDS(here("data/e3_08-04-23.rds")) |> 
    mutate(trainCon=case_when(
    bandOrder=="Original" ~ "800",
    bandOrder=="Reverse" ~ "600",
    TRUE ~ NA_character_
    ), trainCon=as.numeric(trainCon)) 
e3Sbjs <- e3 |> group_by(id,condit,bandOrder) |> summarise(n=n())
testE3 <- e3 |> filter(expMode2 == "Test")
nbins=5
trainE3 <-  e3 |> filter(expMode2=="Train") |> group_by(id,condit,bandOrder, vb) |> 
    mutate(Trial_Bin = cut( gt.train, breaks = seq(1, max(gt.train),length.out=nbins+1),include.lowest = TRUE, labels=FALSE)) 
trainE3_max <- trainE3 |> filter(Trial_Bin == nbins, bandInt==trainCon)

bmtd3 <- readRDS(paste0(here::here("data/model_cache/"), "e3_dist_Cond_Type_RF_2.rds"))

pe3td <- testE3 |> ggplot(aes(x = vb, y = dist,fill=condit,
                               col=ifelse(bandType=="Trained","black","lightgray"),
                               size=ifelse(bandType=="Trained",1.0,.5))) +
  stat_bar + 
  facet_wrap(~bandOrder,ncol=1) +
  theme(legend.title.position="top",
        legend.title = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  labs(x="Band", y="Deviation", fill="Condition") +
  scale_color_manual(values=c("black","lightgray"), 
                     name= "Band Type",
                     labels = c("Trained","Extrapolation"),
                     guide = guide_legend(override.aes = list(fill = NA))
                     ) +
  scale_size_continuous(range = c(0.8, 1.0), guide="none")

pe3ce <- bmtd3 |> emmeans( ~condit *bandOrder*bandType) |>
  gather_emmeans_draws() |>
 condEffects(bandType) + labs(y="Deviation", x="Band Type") + 
 facet_wrap(~bandOrder,ncol=1)

e3_vxBMM <- readRDS(paste0(here::here("data/model_cache", "e3_testVxBand_RF_5k.rds")))

pe3vce <- e3_vxBMM |> emmeans( ~condit* bandOrder* bandInt, 
                       at = list(bandInt = c(100, 350, 600, 800, 1000, 1200))) |>
  gather_emmeans_draws() |> 
  condEffects(bandInt) +
  facet_wrap(~bandOrder,ncol=1) +
  stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
  geom_abline(slope = 1, intercept = 200, linetype = "dashed", color = "gray", size = .5,alpha=.5) +
  scale_x_continuous(breaks = c(100, 350, 600, 800, 1000, 1200), 
                     labels = levels(testE3$vb), 
                     limits = c(100, 1200)) + 
scale_y_continuous(expand=expansion(add=100),breaks=round(seq(300,1400,by=200),2)) +
  theme(legend.title=element_blank()) + 
  labs(y="Velcoity", x="Band")

fe <- fixef(e3_vxBMM)[,1]
fixed_effect_bandInt <- fixef(e3_vxBMM)[,1]["bandInt"]
fixed_effect_interaction1 <- fixef(e3_vxBMM)[,1]["conditVaried:bandInt"]
fixed_effect_interaction2 <- fixef(e3_vxBMM)[,1]["bandOrderReverse:bandInt"]
fixed_effect_interaction3 <- fixef(e3_vxBMM)[,1]["conditVaried:bandOrderReverse:bandInt"]

re <- data.frame(ranef(e3_vxBMM, pars = "bandInt")$id[, ,'bandInt']) |> 
  rownames_to_column("id") |> 
  left_join(e3Sbjs,by="id") |>
  mutate(adjust= fixed_effect_bandInt + fixed_effect_interaction1*(condit=="Varied") + 
           fixed_effect_interaction2*(bandOrder=="Reverse") + 
           fixed_effect_interaction3*(condit=="Varied" & bandOrder=="Reverse"),
  slope = Estimate + adjust )

pid_den3 <- ggplot(re, aes(x = slope, fill = condit)) + 
  geom_density(alpha=.5) + 
  xlim(c(min(re$slope)-.3, max(re$slope)+.3))+
  geom_vline(xintercept = 1, linetype="dashed",alpha=.5) +
   theme(legend.title=element_blank()) + 
  labs(x="Slope Coefficient",y="Density") +
  facet_wrap(~bandOrder,ncol=1)

pid_slopes3 <- re |>  
    mutate(id=reorder(id,slope)) |>
  ggplot(aes(y=id, x=slope,fill=condit,color=condit)) + 
    geom_pointrange(aes(xmin=Q2.5+adjust, xmax=Q97.5+adjust)) + 
    geom_vline(xintercept = 1, linetype="dashed",alpha=.5) +
    theme(legend.title=element_blank(), 
      axis.text.y = element_text(size=6) ) + 
    labs(x="Estimated Slope", y="Participant")  + 
    ggh4x::facet_nested_wrap(bandOrder~condit,axes="all",scales="free_y")

pe3td
pe3ce
pe3vce
pid_den3
pid_slopes3

ggsave(here::here("Presentation/assets/htw_e3_test1.png"), pe3td, width = 4.0, height = 6, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e3_test2.png"), pe3ce, width = 4.0, height = 6, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e3_test3.png"), pe3vce, width = 4.0, height = 6, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e3_test4.png"), pid_den3, width = 4.0, height = 6, units = "in", dpi = 600)
ggsave(here::here("Presentation/assets/htw_e3_test5.png"), pid_slopes3, width = 4.0, height = 6, units = "in", dpi = 600)

```

# Sim Example

@@

# IGAS Task Space

```{r}
#| label: fig-taskSpace
#| fig.cap: A) A visual representation of the combinations of throw parameters (x and y velocities applied to the ball at launch), which resulted in target hits during the testing phase. This empirical solution space was compiled from all of the participants in experiment 2. B) shows the solution space within the context of all of the throws made throughout the testing phase of the experiment.
#| fig-width: 11
#| fig-height: 9

pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,here, RColorBrewer, patchwork)
walk(c("Display_Functions"), ~source(here::here(paste0("Functions/", .x, ".R"))))

e2<- readRDS(here::here('data/igas_e2_cleanedData-final.rds')) %>% mutate(initialVelocityX=X_Velocity,initialVelocityY=Y_Velocity)

taskspace <- e2 %>% filter(AbsDistFromCenter<900)
taskspace$hitOrMiss <- ifelse(taskspace$trialType==11,"Hit Target","Missed Target")

solSpace <- e2 %>% filter(trialType==11)
#solSpace %>% ggplot(aes(x=X_Velocity,y=Y_Velocity)) + geom_point(aes(colour=ThrowPosition),alpha=0.58) + ggtitle("")

solSpace$Result = ifelse(solSpace$ThrowPosition==400,"400",solSpace$ThrowPosition)
solSpace$Result = ifelse(solSpace$ThrowPosition==500,"500",solSpace$Result)
solSpace$Result= ifelse(solSpace$ThrowPosition==625,"625",solSpace$Result)
solSpace$Result = ifelse(solSpace$ThrowPosition==675,"675",solSpace$Result)
solSpace$Result = ifelse(solSpace$ThrowPosition==800,"800",solSpace$Result)
solSpace$Result = ifelse(solSpace$ThrowPosition==900,"900",solSpace$Result)

missSpace <- e2 %>% filter(trialType !=11, abs(X_Velocity)<1000)
missSpace$Result = "Missed Target"
solSpace$Result <- solSpace$Result

# the usual method of changing the legend title does not seem to work after the colours are manually scaled. 
# multiplied velocoties by -1 to make the axes less confusing
ss= solSpace %>% ggplot(aes(x=X_Velocity*-1,y=Y_Velocity*-1)) + 
  geom_point(aes(colour=Result),alpha=0.6) + 
  scale_color_manual(values =brewer.pal(n=6,name="Set1")) +
  labs(color="Throw position:", 
  # title="Empirical Solution Space - Testing Positions",
  # subtitle="The XY velocity coordinates of all the successful throws", 
  x="X Release Velocity", 
  y="Y Release Velocity", )  
  theme(legend.position="top")

fullSpace <- rbind(missSpace,solSpace)

fs<- fullSpace %>% ggplot(aes(x=X_Velocity*-1,y=Y_Velocity*-1,colour=Result)) + 
  geom_point(aes(),alpha=0.6) + scale_color_manual(values =brewer.pal(n=7,name="Set1"))+
  labs(color=element_blank(),
  # title="Full Space of Solution Throws and Missed Throws",
  # subtitle="Colored points represent coordinates of target hits, brown points are coordinates of misses",
  x="X Release Velocity", 
  y="Y Release Velocity") + 
  theme(legend.title = element_blank(), legend.position="top")

ggsave(here::here("Presentation/assets/sol_space.png"),ss)
ggsave(here::here("Presentation/assets/full_space.png"),fs)

```
